Thermoacoustic heat pumps

### 2023-04-22 paper.tex
first draft of standing waves

\documentclass{article}
\usepackage{graphicx} % Required for inserting images

@@

\section{Introduction}
Thermoacoustic heat pumps (TAHP) are a new and upcoming technology. This technology offers an environment-friendly and efficient solution for heating applications.
\newpara{}
In contrast to traditional heating and cooling applications which rely on the vapor compression cycles, TAHP uses the energy stored in sound waves to transfer heat between temperature reservoirs. These mechanisms have a relatively simple design compared to their more-traditional counterparts. TAHP also use more environment-friendly working media such as helium and hydrogen.
Working on these principles, this technology has the potential to offer significant advantages over traditional heating and cooling methods, including reduced energy consumption, lower maintenance costs, and reduced environmental impact.
\newpara{}
This paper will first discuss the physics behind these systems in section X. Section X discusses the applications of TAHP with the difference between standing wave and traveling wave principles in mind. Subsequently, section X discusses the challenges and possible shortcomings of TAHP.\@ This paper will conclude in section X with final remarks.

\section{Working principles}
When a sound wave travels through a tube, gas parcels alternatively compress and expand. This process happens adiabatically. These compressions and expansions change the temperature and pressure of these gas parcels. When the pressure reaches maximum so does the temperature and vice versa. Normally, these temperature differences cancel each other out and the average temperature of the medium stays constant. % FIGURE
A TAHP, however, harnesses these temperature differences in a wave. They do this by using a device called the \emph{stack}. % FIGURE
\newpara{}
The \emph{stack} is designed in such a way that there is a maximum amount of heat transfer from the gas to the stack, but within the stack the heat conduction is limited. The stack should also minimize its obstruction of the gas. A temperature gradient develops on the \emph{stack} when it is placed at a certain location. A heat flow is then generated by placing heat exchangers on both sides of the stack. The reverse of this principle is also true, by applying a temperature gradient to the stack a sound wave can be produced. This system is called a thermoacoustic heat engine or a prime mover, but will not be discussed further in this paper.% FIGURE
\newpara{}
This paper will further discuss the two main types of TAHP, the \emph{standing wave} TAHP and the \emph{traveling wave} TAHP.

\subsection{Standing Waves}
\subsection{Travelling Waves}

@@

\section{Conclusion}

\end{document}

>>>>

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[superscript,biblabel]{cite}

@@

\section{Introduction}
Thermoacoustic heat pumps (TAHP) are a new and upcoming technology. This technology offers an environment-friendly and efficient solution for heating applications.
\newpara{}
In contrast to traditional heating and cooling applications which rely on the vapor compression cycles, TAHP uses the energy stored in sound waves to transfer heat between temperature reservoirs. These mechanisms have a relatively simple design compared to their more-traditional counterparts. TAHP also use more environment-friendly working media such as helium and hydrogen.
Working on these principles, this technology has the potential to offer significant advantages over traditional heating and cooling methods, including reduced energy consumption, lower maintenance costs, and reduced environmental impact.\cite{powerofsound}
\newpara{}
This paper will first discuss the physics behind these systems in section X. Section X discusses the applications of TAHP with the difference between standing wave and traveling wave principles in mind. Subsequently, section X discusses the challenges and possible shortcomings of TAHP.\@ This paper will conclude in section X with final remarks.

\section{Working principles}
\subsection{Standing Waves\cite{powerofsound,enginesandrefrigerators,tijaniLoudSpeaker}}
In a standing wave, gas parcels alternatively compress and expand. This process happens adiabatically. These compressions and expansions change the temperature and pressure of these gas parcels. When the pressure reaches maximum so does the temperature and vice versa.

These temperature variations are relatively small. For sounds at 120 \(dB\), the temperature oscillates up and down by only about 0.02 degrees Celcius. Most air conditioners and refrigerators need to pump heat over a range of 20 degrees or more. So the temperature changes within the gas are too small to be useful.

To handle larger temperature ranges TAHP put the gas in contact with a solid. A solid has a much higher heat capacity per unit volume than gas and can absorb a considerable amount more heat without changing temperature very much. The solid absorbs heat when a gas parcel compresses and moves to the left. The solid will release heat to a gas parcel when it drops in temperature due to expansion and moves to the right. A temperature gradient is created along the solid. The temperature gradient due to this gas parcel is very small, but when multiple of these gas parcels are all undergoing this process, they work as a sort of bucket brigade bringing heat from one side to the other. % FIGURE
This cycle for a gas parcel plotted in a \(p-V\) diagram shows that there is work absorbed by the gas and is typical for a heat pump.  % FIGURE
This process can be reversed where a temperature gradient along the solid is provided by an outside source which is then converted into acoustic power. Such a device is called a \emph{thermoacoustic heat engine} or \emph{primer mover}.

The solid used in TAHP should have good thermal contact when the parcel is stationary but poor thermal contact when the parcel is moving. The solid should not conduct heat from one end to the other and thus have a low conductivity. The quest of finding this material is a large part of making TAHP commercially viable, but plastic is often sufficient for experimental applications.

TAHP use a stack of plates of the solid previously discussed called the \emph{stack}. These plates are mounted parallel to each other. Detailed analysis shows that the optimal distance between these plates is about four times the \emph{thermal penetration depth} \(\delta_k = \sqrt{\kappa/\pi f \rho c_p}\). % DISCUSS THIS FURTHER

The most rudimentary form of a TAHP consists of a resonator, gas, stack, acoustic transducer, and hot and cold heat exchangers. The acoustic transducer applies acoustic power \(W\) to the gas in the resonator such that a standing wave of the first harmonic is formed in the resonator. A stack is located at about a fourth of the resonator length from the transducer. A hot heat exchanger is mounted on the left at the side of the stack which is closest to the node where temperature and pressure will be at their maximum. A cold heat exchanger is then mounted on the right side closest to the anti-node.

The gasses used in TAHP are typically noble gasses such as helium, neon, argon, or a mixture of these gases. These gasses are used because they have low molecular weights and are highly compressible, which makes them ideal for use in thermoacoustic refrigerators. Additionally, they have low thermal conductivity, which helps to minimize heat transfer between the hot and cold sides of the device.

The bucket brigade of gas parcels pumps heat from the cold end of the stack to the hot end and creates a temperature gradient. In the case of a \emph{thermoacoustic refrigerator}, the cold exchanger is connected to a room to be cooled at \(T_c\), and the hot exchanger is connected to a heat sink at \(T_h\). This is reversed when the desired effect is the heating of a room. This technology is, however, the most promising in the former case.

The efficiency for heat pumps, known as the \emph{coefficient of performance (COP)}, is limited by the second law of thermodynamics and is defined by the Carnot efficiency \(T_c/(T_h-T_c)\)
The efficiency of this rudimentary device is, sadly, disappointing.

% LOSSES
% EFFICIENCY
% VARIATIONS

% - COP
% - reasonable efficiency and extreme
% tube & speaker -> resonance
% Helium (good heat transfer properties)
\subsection{Travelling Waves}

@@

\section{Conclusion}
\bibliography{uni}
\bibliographystyle{unsrt}
\end{document}
